---
title: "MAP Scenario Dashboards"
output:
  html_document:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 13,
  fig.height = 7.5,
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)
```

```{r load}
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(knitr)

theme_map_scenario <- function() {
  ggplot2::theme_minimal(base_size = 9) +
    ggplot2::theme(
      plot.title = ggplot2::element_text(size = 8, face = "bold"),
      plot.subtitle = ggplot2::element_text(size = 7),
      axis.title = ggplot2::element_text(size = 7),
      axis.text = ggplot2::element_text(size = 7),
      legend.title = ggplot2::element_text(size = 7),
      legend.text = ggplot2::element_text(size = 7),
      strip.text = ggplot2::element_text(size = 7)
    )
}

theme_set(theme_map_scenario())

repo_root <- local({
  candidates <- unique(normalizePath(
    c(getwd(), file.path(getwd(), ".."), file.path(getwd(), "..", "..")),
    mustWork = FALSE
  ))
  hit <- candidates[
    file.exists(file.path(candidates, "code", "in-vitro-utils", "io.R"))
  ][1]
  if (is.na(hit) || !nzchar(hit)) {
    stop("Could not locate repository root from working directory: ", getwd())
  }
  hit
})

source(file.path(repo_root, "code", "in-vitro-utils", "io.R"))
source(file.path(repo_root, "code", "in-vitro-utils", "runner.R"))

ivt_source_map_model(repo_root)
```

This report rebuilds the qualitative scenario dashboards from `ploidyOxygen/workflow.Rmd`, but runs them through the updated O2 supply-demand MAP model with the fixed-oxygen in vitro settings.

Translation strategy:

- hardcode one small default MAP parameter set near the top of the notebook, with provenance comments
- keep only the old buffering-model parameters that have a clear MAP analogue
- choose a few MAP-only parameter values that numerically suppress the direct ploidy dependence of hypoxia response without changing the MAP switches
- preserve the old dashboard starting states as exact chromosome-count starts at 44 and 88

```{r params}
cfg <- ivt_build_default_cfg(
  repo_root = repo_root,
  dt = 0.1,
  init_total_size = 1e6,
  o2_upper_bound = 21,
  fixed_oxygen = TRUE
)

scenario_cfg <- cfg
scenario_cfg$Crowding <- FALSE
scenario_cfg$K <- 1e30
scenario_cfg$init_total_size <- 1
scenario_cfg$o2_Nref <- scenario_cfg$init_total_size

legacy_theta <- list(
  p_misseg = 0.01,
  p_wgd = 0.00001,
  beta = 2.5,
  n_exp = 10,
  smax = 0.8,
  k_o = 3.670835,
  k_o_mis = 0.1,
  lam_max = 1.138527,
  lam_min = 0.482247,
  mu_2N = 41.34470,
  sd_2N = 1.410775,
  mu_4N = 55.69906,
  sd_4N = 16.30977
)

# Hardcoded defaults for dashboard tinkering.
# Provenance:
# - lambda / WGD / missegregation values come from the old qualitative buffering dashboard.
# - MAP-only terms keep the same "weak direct ploidy dependence" translation that this report
#   previously used, but are now written out explicitly so the active scenario is visible in one place.
# - start from the in vitro parameter-table defaults so required MAP housekeeping terms like `tau_O2`
#   are present, then overwrite only the handful of scenario-facing values we want to tinker with here.
# - 2N / 4N starts are exact point masses at N = 44 and N = 88; the IC means below are only
#   metadata/reference values and are not used unless you switch away from point-mass starts.
scenario_run_params <- ivt_load_default_run_params(scenario_cfg)

# Oxygen-only lower division-rate asymptote.
scenario_run_params$lam_min <- legacy_theta$lam_min
# Oxygen-only upper division-rate asymptote.
scenario_run_params$lam_max <- legacy_theta$lam_max
# Oxygen half-saturation in lambda(O2).
scenario_run_params$k_o <- legacy_theta$k_o
# Whole-genome doubling probability per division.
scenario_run_params$p_wgd <- 0.001#legacy_theta$p_wgd
# Small oxygen-independent missegregation floor kept in MAP.
scenario_run_params$p_mis_base <- legacy_theta$p_misseg * (1 - 21 / (21 + legacy_theta$k_o_mis))
# Oxygen-sensitive missegregation increment on top of the floor.
scenario_run_params$p_misseg <- legacy_theta$p_misseg - scenario_run_params$p_mis_base
# Very small half-saturation to keep low-O2 missegregation near the old ceiling.
scenario_run_params$k_o_mis <- 1e-05
# Strength of nullisomy-risk death on missegregation loss branches.
scenario_run_params$gamma_loss <- 0.1
# Turn off direct ploidy modulation of growth under hypoxia.
scenario_run_params$alpha_o2 <- 0
# Keep hypoxia-linked death present but weak.
scenario_run_params$mu_hp <- 0.001
# Ploidy exponent on hypoxia-linked death.
scenario_run_params$gamma_mu <- 1
# Oxygen threshold for the hypoxia-response switch.
scenario_run_params$O2_crit <- 1
# Steepness of the hypoxia-response switch.
scenario_run_params$n_O <- 5
# Reference 2N initial-condition mean; only used if you stop using point-mass starts.
scenario_run_params$init_mean_2N <- legacy_theta$mu_2N
# Reference 2N initial-condition SD; only used if you stop using point-mass starts.
scenario_run_params$init_sd_2N <- legacy_theta$sd_2N
# Reference 4N initial-condition mean; only used if you stop using point-mass starts.
scenario_run_params$init_mean_4N <- legacy_theta$mu_4N
# Reference 4N initial-condition SD; only used if you stop using point-mass starts.
scenario_run_params$init_sd_4N <- legacy_theta$sd_4N

# This canonicalizes a few shared MAP fields and fills housekeeping defaults expected by the C++ simulator.
scenario_run_params <- normalize_run_params_common(scenario_run_params, cfg = scenario_cfg)

active_run_params_table <- tibble::tibble(
  parameter = sort(names(scenario_run_params)),
  value = vapply(scenario_run_params[sort(names(scenario_run_params))], function(x) {
    if (length(x) == 0L || is.null(x)) return(NA_character_)
    if (is.numeric(x[[1]])) return(as.character(signif(as.numeric(x[[1]]), 6)))
    as.character(x[[1]])
  }, character(1))
)

parameter_provenance <- tibble::tribble(
  ~parameter, ~value, ~provenance,
  "lam_min", scenario_run_params$lam_min, "Old buffering dashboard default.",
  "lam_max", scenario_run_params$lam_max, "Old buffering dashboard default.",
  "k_o", scenario_run_params$k_o, "Old buffering dashboard default.",
  "p_wgd", scenario_run_params$p_wgd, "Old buffering dashboard default.",
  "p_mis_base", scenario_run_params$p_mis_base, "Derived from old buffering `p_misseg` at high oxygen to create a small oxygen-independent MAP floor.",
  "p_misseg", scenario_run_params$p_misseg, "Derived from old buffering `p_misseg` after separating the MAP floor into `p_mis_base`.",
  "k_o_mis", scenario_run_params$k_o_mis, "Small ad hoc MAP value to keep low-O2 missegregation near the old ceiling.",
  "gamma_loss", scenario_run_params$gamma_loss, "Explicitly set here for dashboard tinkering; controls nullisomy-risk death on loss branches.",
  "alpha_o2", scenario_run_params$alpha_o2, "Set to zero so growth stays effectively O2-only.",
  "mu_hp", scenario_run_params$mu_hp, "Small ad hoc MAP value to keep hypoxia-linked death weak.",
  "gamma_mu", scenario_run_params$gamma_mu, "Simple neutral MAP setting retained from the prior translation.",
  "O2_crit", scenario_run_params$O2_crit, "Ad hoc MAP setting used in the prior translation.",
  "n_O", scenario_run_params$n_O, "Ad hoc MAP setting used in the prior translation.",
  "init_mean_2N", scenario_run_params$init_mean_2N, "Old buffering fitted IC mean; reference only for this notebook.",
  "init_sd_2N", scenario_run_params$init_sd_2N, "Old buffering fitted IC SD; reference only for this notebook.",
  "init_mean_4N", scenario_run_params$init_mean_4N, "Old buffering fitted IC mean; reference only for this notebook.",
  "init_sd_4N", scenario_run_params$init_sd_4N, "Old buffering fitted IC SD; reference only for this notebook."
)

param_translation <- tibble::tribble(
  ~legacy_parameter, ~legacy_value, ~map_value_used, ~status, ~note,
  "lam_min", legacy_theta$lam_min, scenario_run_params$lam_min, "direct", "Same O2-only lower growth asymptote.",
  "lam_max", legacy_theta$lam_max, scenario_run_params$lam_max, "direct", "Same O2-only upper growth asymptote.",
  "k_o", legacy_theta$k_o, scenario_run_params$k_o, "direct", "Same half-saturation term in lambda(O2).",
  "p_misseg", legacy_theta$p_misseg, scenario_run_params$p_misseg, "approximate", "Split into a tiny MAP baseline floor plus a near-saturating stress-linked increment so the 0% and 21% oxygen cases qualitatively match the old curve.",
  "k_o_mis", legacy_theta$k_o_mis, scenario_run_params$k_o_mis, "approximate", "Not directly comparable; MAP uses a death-rate half-saturation, so a very small value is used to keep low-O2 missegregation near the old ceiling while preserving weak ploidy sensitivity.",
  "gamma_loss", NA_real_, scenario_run_params$gamma_loss, "map_only", "Controls nullisomy-risk death on loss branches; no direct old buffering-model analogue.",
  "p_wgd", legacy_theta$p_wgd, scenario_run_params$p_wgd, "direct", "Same whole-genome doubling probability.",
  "beta", legacy_theta$beta, NA_real_, "approximate", "No 1:1 MAP equivalent; the old ploidy-buffering effect is only approximated by keeping hypoxia-linked death numerically weak rather than structurally removing it.",
  "n_exp", legacy_theta$n_exp, NA_real_, "approximate", "No 1:1 MAP equivalent; the old sharp ploidy-buffering curvature is not reproduced explicitly in MAP.",
  "smax", legacy_theta$smax, NA_real_, "approximate", "No direct survival cap in MAP; qualitative correspondence is achieved by making the MAP hypoxia-death module very weak rather than by matching smax itself.",
  "mu_2N", legacy_theta$mu_2N, 44, "approximate", "Old dashboard already imposed an exact 2N scenario start, so the MAP scenario starts at N = 44 rather than the old fitted IC mean.",
  "sd_2N", legacy_theta$sd_2N, scenario_run_params$init_sd_2N, "approximate", "Recorded for reference, but scenario starts use an exact point mass at N = 44.",
  "mu_4N", legacy_theta$mu_4N, 88, "approximate", "Old dashboard already imposed an exact 4N scenario start, so the MAP scenario starts at N = 88 rather than the old fitted IC mean.",
  "sd_4N", legacy_theta$sd_4N, scenario_run_params$init_sd_4N, "approximate", "Recorded for reference, but scenario starts use an exact point mass at N = 88."
)

map_specific_terms <- tibble::tribble(
  ~map_parameter, ~value_used, ~role,
  "p_mis_base", scenario_run_params$p_mis_base, "Set to the old model's high-oxygen residual missegregation rate so MAP still has a small oxygen-independent floor.",
  "gamma_loss", scenario_run_params$gamma_loss, "Controls how strongly nullisomy risk reduces survival on missegregation loss branches.",
  "alpha_o2", scenario_run_params$alpha_o2, "Set to zero so the MAP growth module keeps the old O2-only lambda(O2) shape even though O2_growth stays enabled.",
  "gamma_growth", scenario_run_params$gamma_growth, "Retained but inert because alpha_o2 is zero.",
  "mu_hp", scenario_run_params$mu_hp, "Set small so hypoxia-linked death stays present but weak.",
  "gamma_mu", scenario_run_params$gamma_mu, "Kept simple; residual ploidy sensitivity persists, but only through the weak mu_hp term.",
  "O2_crit", scenario_run_params$O2_crit, "Set low so the hypoxia-response weight is near one at 0% and near zero at high oxygen.",
  "n_O", scenario_run_params$n_O, "Set high enough to make the hypoxia-response drop steeply with oxygen.",
  "gamma_loss", scenario_run_params$gamma_loss, "Retained from the active MAP workflow because it governs intrinsic missegregation-loss survival rather than the direct O2 response."
)

base_scenario <- list(
  label = "base scenario",
  o2_requested = 0,
  start_N = 44,
  cohort = "2N",
  duration_days = 500,
  times = seq(0, 500, length.out = 5),
  init_total_size = 1
)

make_scenario <- function(base, label, o2_requested = NULL, start_N = NULL, cohort = NULL, duration_days = NULL, times = NULL) {
  out <- base
  if (!is.null(label)) out$label <- label
  if (!is.null(o2_requested)) out$o2_requested <- o2_requested
  if (!is.null(start_N)) out$start_N <- start_N
  if (!is.null(cohort)) out$cohort <- cohort
  if (!is.null(duration_days)) out$duration_days <- duration_days
  if (!is.null(times)) out$times <- times
  out
}
```

```{r source-note, results='asis'}
cat(
  paste0(
    "This notebook now uses a hardcoded dashboard parameter set defined near the top of the file. ",
    "The intent is to make the scenario fully inspectable and easy to edit, rather than to automatically pull the latest fitted MAP parameter set."
  )
)
```

```{r helper-fns}
point_mass_initial_state <- function(grid_pre, start_N) {
  idx <- which.min(abs(as.numeric(grid_pre) - as.numeric(start_N)))
  out <- rep(0, length(grid_pre))
  out[idx] <- 1
  out
}

central_log_growth <- function(time, live_cells) {
  time <- as.numeric(time)
  live_cells <- pmax(as.numeric(live_cells), 1e-12)
  out <- rep(NA_real_, length(time))
  for (i in seq_along(time)) {
    if (length(time) == 1L) {
      out[i] <- NA_real_
    } else if (i == 1L) {
      dt <- time[2] - time[1]
      out[i] <- if (dt > 0) (log(live_cells[2]) - log(live_cells[1])) / dt else NA_real_
    } else if (i == length(time)) {
      dt <- time[i] - time[i - 1L]
      out[i] <- if (dt > 0) (log(live_cells[i]) - log(live_cells[i - 1L])) / dt else NA_real_
    } else {
      dt <- time[i + 1L] - time[i - 1L]
      out[i] <- if (dt > 0) (log(live_cells[i + 1L]) - log(live_cells[i - 1L])) / dt else NA_real_
    }
  }
  out
}

simulate_map_scenario <- function(o2_requested,
                                  start_N,
                                  cohort,
                                  times,
                                  cfg,
                                  run_params,
                                  model_core = NULL) {
  if (is.null(model_core)) model_core <- build_model_core(cfg = cfg)
  duration_days <- max(times)
  actual_o2 <- max(0, min(cfg$o2_S0_upper_bound, as.numeric(o2_requested)))
  obs_days_local <- seq(0, duration_days, by = cfg$DT)
  obs_steps_cpp <- as.integer(round(obs_days_local / cfg$DT))
  init_state <- point_mass_initial_state(model_core$grid_pre, start_N) * cfg$init_total_size
  vol_by_N <- cell_volume_mm3_by_N(model_core$grid_pre, run_params = run_params, cfg = cfg)

  sim_cpp <- cpp_o2simps_simulate_one(
    init_state = as.numeric(init_state),
    N0min = as.integer(cfg$N_MIN),
    N0max = as.integer(cfg$N_MAX),
    N1min = as.integer(cfg$N_MIN),
    N1max = as.integer(cfg$N_MAX),
    obs_steps = as.integer(obs_steps_cpp),
    sim_end_step = as.integer(max(obs_steps_cpp)),
    DT = as.numeric(cfg$DT),
    dose = 0,
    dose_ref = as.numeric(cfg$dose_ref),
    treat_day = Inf,
    fit_treatment = FALSE,
    alpha = as.numeric(.first_non_null_local(run_params$alpha, 0)),
    gamma = as.numeric(.first_non_null_local(run_params$gamma, 1)),
    tx_mult_min = as.numeric(cfg$tx_mult_min),
    crowding_enabled = isTRUE(cfg$Crowding),
    crowding = as.character(cfg$crowding),
    K = as.numeric(cfg$K),
    min_pop = as.numeric(cfg$min_pop),
    O2_crit = as.numeric(.first_non_null_local(run_params$O2_crit, cfg$o2_crit_init, 1.0)),
    o2_feedback = FALSE,
    o2_S0 = as.numeric(actual_o2),
    kappa_O = as.numeric(.first_non_null_local(run_params$kappa_O, 1.0)),
    tau_O2 = as.numeric(.first_non_null_local(run_params$tau_O2, cfg$tau_O2, cfg$tau_O2_init, 2.0)),
    o2_Nref = as.numeric(cfg$o2_Nref),
    o2_min = as.numeric(actual_o2),
    eta_o2 = as.numeric(.first_non_null_local(run_params$eta_o2, 1.0)),
    o2_cache_bin_pct = 0.01,
    o2_cache_hysteresis_pct = 0.005,
    o2_cache_profile = FALSE,
    O2_growth = isTRUE(cfg$O2_growth),
    lam_min = as.numeric(run_params$lam_min),
    lam_max = as.numeric(run_params$lam_max),
    k_o = as.numeric(run_params$k_o),
    has_p_misseg = !is.null(run_params$p_misseg),
    p_mis_base = as.numeric(.first_non_null_local(run_params$p_mis_base, 1e-5)),
    p_misseg = as.numeric(.first_non_null_local(run_params$p_misseg, 0)),
    k_o_mis = as.numeric(.first_non_null_local(run_params$k_o_mis, 50.0)),
    has_pmis_endpoints = FALSE,
    pmis_O2_0 = 0,
    pmis_O2_1 = 0,
    p_const = 0,
    p_wgd = as.numeric(.first_non_null_local(run_params$p_wgd, 0)),
    boundary = as.character(.first_non_null_local(run_params$boundary, "drop")),
    eps_tail = 1e-8,
    gamma_loss = as.numeric(.first_non_null_local(run_params$gamma_loss, 0.1)),
    misseg_loss_survival = as.character(.first_non_null_local(run_params$misseg_loss_survival, cfg$misseg_loss_survival, "nullisomy")),
    buffer_smax = as.numeric(.first_non_null_local(run_params$buffer_smax, cfg$buffer_smax_init, 1.0)),
    buffer_beta = as.numeric(.first_non_null_local(run_params$buffer_beta, cfg$buffer_beta_init, 0.0)),
    buffer_n_exp = as.numeric(.first_non_null_local(run_params$buffer_n_exp, cfg$buffer_n_exp_init, 1.0)),
    N_unit = as.integer(cfg$N_UNIT),
    beta_size = as.numeric(.first_non_null_local(run_params$beta_size, default_beta_size_prior_center())),
    alpha_o2 = as.numeric(.first_non_null_local(run_params$alpha_o2, cfg$alpha_o2_init, 0.5)),
    gamma_growth = as.numeric(.first_non_null_local(run_params$gamma_growth, cfg$gamma_growth_init, 2.0)),
    mu_hp = as.numeric(.first_non_null_local(run_params$mu_hp, cfg$mu_hp_init, 1e-3)),
    gamma_mu = as.numeric(.first_non_null_local(run_params$gamma_mu, cfg$gamma_mu_init, 1.0)),
    n_O = as.numeric(.first_non_null_local(run_params$n_O, cfg$n_O_init, 1.0)),
    ploidy_O2_death = assert_canonical_ploidy_o2_death_mode(
      .first_non_null_local(cfg$ploidy_O2_death, run_params$ploidy_O2_death, "uniform")
    ),
    k_clear = as.numeric(.first_non_null_local(run_params$k_clear, cfg$k_clear_init, 1e-3)),
    vol_by_N = as.numeric(vol_by_N),
    burden_floor = as.numeric(cfg$burden_log_eps),
    return_full_trajectory = TRUE
  )

  live_state_obs <- as.matrix(sim_cpp$live_state_obs)
  live_cells <- as.numeric(sim_cpp$Ntot_live_obs)
  growth_df <- tibble::tibble(
    day = obs_days_local,
    live_cells = live_cells,
    growth_rate = central_log_growth(obs_days_local, live_cells)
  )

  snapshot_df <- dplyr::bind_rows(lapply(times, function(t_req) {
    idx <- which.min(abs(obs_days_local - t_req))
    w <- as.numeric(live_state_obs[idx, ])
    s <- sum(w)
    tibble::tibble(
      snapshot_time = factor(
        paste0("t=", sprintf("%.1f", round(t_req, 1))),
        levels = paste0("t=", sprintf("%.1f", round(times, 1)))
      ),
      snapshot_day_requested = as.numeric(t_req),
      snapshot_day_simulated = as.numeric(obs_days_local[idx]),
      N = as.numeric(model_core$grid_pre),
      fraction = if (s > 0) w / s else 0
    )
  }))

  list(
    o2_requested = as.numeric(o2_requested),
    oxygen_pct = actual_o2,
    start_N = as.numeric(start_N),
    cohort = cohort,
    growth_df = growth_df,
    snapshot_df = snapshot_df,
    o2_trace_df = tibble::tibble(
      day = obs_days_local,
      oxygen_target = as.numeric(.first_non_null_local(sim_cpp$O2_target_obs, rep(actual_o2, length(obs_days_local)))),
      oxygen_effective = as.numeric(.first_non_null_local(sim_cpp$O2_eff_obs, rep(actual_o2, length(obs_days_local))))
    ),
    model_core = model_core
  )
}

plot_map_dashboard <- function(scenario, run_params) {
  O_seq <- seq(0, 21, length.out = 200)
  N_seq <- seq(scenario$model_core$grid_pre[[1]], max(scenario$model_core$grid_pre), by = 1)

  death_df <- tibble::tibble(
    N = N_seq,
    mu_eff = .mu_eff_of_O2(
      O2 = scenario$oxygen_pct,
      run_params = run_params,
      N = N_seq
    )
  )

  p1 <- ggplot(death_df, aes(N, mu_eff)) +
    geom_line(color = "#274c77", linewidth = 1) +
    geom_vline(xintercept = scenario$start_N, linetype = "dashed", color = "#c1121f") +
    labs(
      title = "Hypoxia death penalty by chromosome count",
      subtitle = paste0("mu_eff(N) at fixed O2 = ", signif(scenario$oxygen_pct, 3), "%"),
      x = "Chromosome-count state N",
      y = "Death rate"
    )

  pmis_df <- tibble::tibble(
    O2 = O_seq,
    p_misseg = .pmisseg_of_O2(
      O2 = O_seq,
      run_params = run_params,
      N = scenario$start_N
    )
  )

  p2 <- ggplot(pmis_df, aes(O2, p_misseg)) +
    geom_line(color = "#bc6c25", linewidth = 1) +
    geom_vline(xintercept = scenario$oxygen_pct, linetype = "dotted", color = "#c1121f") +
    geom_point(
      data = tibble::tibble(
        O2 = scenario$oxygen_pct,
        p_misseg = .pmisseg_of_O2(scenario$oxygen_pct, run_params = run_params, N = scenario$start_N)
      ),
      aes(O2, p_misseg),
      color = "#c1121f",
      size = 2
    ) +
    labs(
      title = "Missegregation versus oxygen",
      subtitle = paste0("Evaluated at start state N = ", scenario$start_N),
      x = "Oxygen (%)",
      y = "Per-division missegregation probability"
    )

  lambda_df <- tibble::tibble(
    O2 = O_seq,
    lambda = growth_lambda(
      O2 = O_seq,
      N = scenario$start_N,
      lam_min = run_params$lam_min,
      lam_max = run_params$lam_max,
      k_o = run_params$k_o
    )
  )

  p3 <- ggplot(lambda_df, aes(O2, lambda)) +
    geom_line(color = "#6a4c93", linewidth = 1) +
    geom_vline(xintercept = scenario$oxygen_pct, linetype = "dotted", color = "#c1121f") +
    geom_point(
      data = tibble::tibble(
        O2 = scenario$oxygen_pct,
        lambda = growth_lambda(scenario$oxygen_pct, scenario$start_N, run_params$lam_min, run_params$lam_max, run_params$k_o)
      ),
      aes(O2, lambda),
      color = "#c1121f",
      size = 2
    ) +
    labs(
      title = "Cycle rate versus oxygen",
      subtitle = "Direct carry-over from the old O2-only lambda(O2) term",
      x = "Oxygen (%)",
      y = "Division rate"
    )

  p4 <- ggplot(scenario$growth_df, aes(day, growth_rate)) +
    geom_line(color = "#2a9d8f", linewidth = 0.9) +
    geom_point(
      data = scenario$growth_df |> dplyr::filter(day %in% as.numeric(gsub("^t=", "", levels(scenario$snapshot_df$snapshot_time)))),
      size = 1.8,
      color = "#1d3557"
    ) +
    labs(
      title = "Net growth over time",
      subtitle = "Estimated from one direct simulator trajectory; no passage/segment reseeding in this notebook path",
      x = "Time (days)",
      y = "Net growth rate"
    )

  loss_sizes <- c(1L, 2L, 3L)
  loss_mode <- as.character(.first_non_null_local(run_params$misseg_loss_survival, cfg$misseg_loss_survival, "nullisomy"))
  loss_df <- dplyr::bind_rows(lapply(loss_sizes, function(m_loss_i) {
    survival <- if (identical(loss_mode, "buffering")) {
      sN <- as.numeric(.first_non_null_local(run_params$buffer_smax, 1.0)) *
        exp(-as.numeric(.first_non_null_local(run_params$buffer_beta, 0.0)) *
          ((2 * as.integer(.first_non_null_local(run_params$N_unit, 22L))) / N_seq)^
            as.numeric(.first_non_null_local(run_params$buffer_n_exp, 1.0)))
      pmax(0, pmin(1, sN))^m_loss_i
    } else {
      .loss_survival_nullisomy(
        N = N_seq,
        m_loss = m_loss_i,
        gamma_loss = as.numeric(.first_non_null_local(run_params$gamma_loss, 0.1)),
        N_unit = as.integer(.first_non_null_local(run_params$N_unit, 22L))
      )
    }
    tibble::tibble(
      N = N_seq,
      survival = survival,
      loss_size = factor(
        paste0("loss=", m_loss_i),
        levels = paste0("loss=", loss_sizes)
      )
    )
  }))
  loss_subtitle <- if (identical(loss_mode, "buffering")) {
    "Symmetric buffering survival modifier for gain/loss missegregation events"
  } else {
    "Nullisomy-risk survival modifier for one-step loss events at current gamma_loss"
  }

  p6 <- ggplot(loss_df, aes(N, survival, color = loss_size)) +
    geom_line(linewidth = 0.9) +
    geom_vline(xintercept = scenario$start_N, linetype = "dashed", color = "#c1121f") +
    labs(
      title = "Missegregation survival by chromosome count",
      subtitle = loss_subtitle,
      x = "Chromosome-count state N",
      y = "Survival after missegregation event",
      color = "Loss size"
    ) +
    theme(legend.position = "top")

  p5 <- ggplot(scenario$snapshot_df, aes(N, fraction, color = snapshot_time, fill = snapshot_time)) +
    geom_area(alpha = 0.2, position = "identity") +
    geom_line(linewidth = 0.9) +
    scale_y_sqrt() +
    labs(
      title = "Ploidy distribution snapshots",
      subtitle = paste0(
        "Requested O2 = ", signif(scenario$o2_requested, 3), "%; simulated O2 = ", signif(scenario$oxygen_pct, 3), "%"
      ),
      x = "Chromosome-count state N",
      y = "Fraction",
      color = "Time",
      fill = "Time"
    ) +
    theme(legend.position = "top")

  gridExtra::grid.arrange(
    grobs = list(p1, p2, p3, p4, p5, p6),
    layout_matrix = rbind(
      c(1, 6),
      c(2, 5),
      c(3, 5),
      c(4, 5)
    ),
    widths = c(1.0, 1.25)
  )
}

model_core <- build_model_core(cfg = scenario_cfg)
scenario_specs <- list(
  make_scenario(
    base_scenario,
    label = "0% oxygen, 4N start",
    o2_requested = 0,
    start_N = 88,
    cohort = "4N",
    times = seq(1, 501, length.out = 5)
  ),
  make_scenario(
    base_scenario,
    label = "0% oxygen, 2N start",
    o2_requested = 0,
    start_N = 44,
    cohort = "2N",
    times = seq(1, 100, length.out = 5)
  ),
  make_scenario(
    base_scenario,
    label = "21% oxygen, 4N start",
    o2_requested = 21,
    start_N = 88,
    cohort = "4N"
  ),
  make_scenario(
    base_scenario,
    label = "21% oxygen, 2N start",
    o2_requested = 21,
    start_N = 44,
    cohort = "2N"
  )
)

scenario_runs <- lapply(seq_along(scenario_specs), function(i) {
  spec <- scenario_specs[[i]]
  simulate_map_scenario(
    o2_requested = spec$o2_requested,
    start_N = spec$start_N,
    cohort = spec$cohort,
    times = spec$times,
    cfg = scenario_cfg,
    run_params = scenario_run_params,
    model_core = model_core
  )
})
names(scenario_runs) <- c("dashboard_0_4N", "dashboard_0_2N", "dashboard_22_4N", "dashboard_22_2N")
```

## Parameter Translation

```{r parameter-provenance}
knitr::kable(
  parameter_provenance |>
    mutate(value = signif(value, 5)),
  caption = "Hardcoded dashboard parameters and their provenance."
)
```

```{r active-params}
knitr::kable(
  active_run_params_table,
  caption = "Full parameter object passed to the model after normalization/default-filling."
)
```

```{r translation-table}
knitr::kable(
  param_translation |>
    mutate(
      legacy_value = signif(legacy_value, 5),
      map_value_used = ifelse(is.na(map_value_used), NA, signif(map_value_used, 5))
    ),
  caption = "Legacy buffering-scenario parameters mapped into the MAP report."
)
```

```{r map-specific-table}
knitr::kable(
  map_specific_terms |>
    mutate(value_used = signif(value_used, 5)),
  caption = "MAP-only terms inherited from the current in vitro workflow."
)
```

The main non-trivial translation choices are:

- `beta`, `n_exp`, and `smax` do not have a clean one-parameter MAP equivalent, so the qualitative buffering effect is only approximated rather than matched exactly.
- `alpha_o2` is set to zero and `mu_hp` is set small, which keeps the MAP switches intact but numerically suppresses most of the direct ploidy dependence of hypoxia response.
- `p_mis_base`, `p_misseg`, `k_o_mis`, `O2_crit`, and `n_O` are chosen together to mimic the old low-O2 versus high-O2 missegregation contrast, because MAP does not use the old direct `O2 / (O2 + k_o_mis)` form.
- the high-oxygen dashboards in this notebook are now stated directly as `21%` oxygen, because the fixed-oxygen MAP workflow clips oxygen to the `0-21%` range anyway
- the old dashboards were already defined by exact starts at `N = 44` and `N = 88`, so the MAP scenarios keep those starts explicitly instead of trying to force the old `mu_*` values into the new workflow

## 0% Oxygen, 4N Start

```{r dashboard-0-4N}
plot_map_dashboard(scenario_runs$dashboard_0_4N, scenario_run_params)
```

## 0% Oxygen, 2N Start

```{r dashboard-0-2N}
plot_map_dashboard(scenario_runs$dashboard_0_2N, scenario_run_params)
```

## 21% Oxygen, 4N Start

```{r dashboard-22-4N}
plot_map_dashboard(scenario_runs$dashboard_22_4N, scenario_run_params)
```

## 21% Oxygen, 2N Start

```{r dashboard-22-2N}
plot_map_dashboard(scenario_runs$dashboard_22_2N, scenario_run_params)
```
