---
title: "In Vitro Adapter Demonstration"
output:
  html_document:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)
```

```{r load}
library(ggplot2)
library(dplyr)
library(tidyr)

repo_root <- local({
  candidates <- unique(normalizePath(
    c(getwd(), file.path(getwd(), ".."), file.path(getwd(), "..", "..")),
    mustWork = FALSE
  ))
  hit <- candidates[
    file.exists(file.path(candidates, "code", "in-vitro-utils", "io.R"))
  ][1]
  if (is.na(hit) || !nzchar(hit)) {
    stop("Could not locate repository root from working directory: ", getwd())
  }
  hit
})

source(file.path(repo_root, "code", "in-vitro-utils", "io.R"))
source(file.path(repo_root, "code", "in-vitro-utils", "lineage_adapter.R"))
source(file.path(repo_root, "code", "in-vitro-utils", "runner.R"))
source(file.path(repo_root, "code", "in-vitro-utils", "summaries.R"))
source(file.path(repo_root, "code", "in-vitro-utils", "plotting.R"))
source(file.path(repo_root, "code", "in-vitro-utils", "objective.R"))

ivt_source_map_model(repo_root)
```

This document demonstrates the adapter layer for the in vitro lineage data.

What it does:

- loads `fit_data.Rds` and `jobs_2N.Rds`
- chooses one terminal 2N lineage automatically for the single-lineage demo plots
- converts that lineage into a sequence of simulation segments
- loads the best optimized parameter set from `workflow/sampled_objective_draws`
- evaluates the full multi-lineage objective across all `jobs_2N` and `jobs_4N` segments
- if a row has multiple `data_ids`, aggregates passage metadata across them by median
- simulates each passage for its recorded duration
- reseeds the next passage to its recorded initial cell count using the final ploidy fractions from the parent passage
- returns outputs aligned back to the original passage IDs

```{r build-adapter}
ivt_objects <- ivt_load_fit_objects(repo_root)
fallback_max_passage_days <- 14

terminal_key <- ivt_choose_demo_terminal_key(
  jobs = ivt_objects$jobs_2N,
  fit_data = ivt_objects$fit_data,
  min_kary = 1L
)

adapter <- ivt_build_lineage_adapter(
  jobs = ivt_objects$jobs_2N,
  fit_data = ivt_objects$fit_data,
  terminal_sim_key = terminal_key,
  cohort = "2N",
  max_segment_days = fallback_max_passage_days
)
```

Chosen lineage: `r terminal_key`

Assumptions used here:

- one `jobs_2N` row is treated as one simulation segment
- if multiple `data_ids` attach to one row, `passage_duration`, `initial_cells`, and `final_cells` are aggregated by median
- fallback maximum passage length if duration is missing = `r fallback_max_passage_days` days

```{r run-lineage}
cfg <- ivt_build_default_cfg(
  repo_root = repo_root,
  dt = 0.1,
  init_total_size = 1e6,
  o2_upper_bound = 21,
  fixed_oxygen = TRUE
)

best_sample_path <- file.path(repo_root, "workflow", "sampled_objective_draws", "invitro_objective_array.tsv")
best_fit <- tryCatch(
  ivt_load_best_sampled_run_params(cfg = cfg, repo_root = repo_root, tsv_path = best_sample_path),
  error = function(e) {
    list(
      run_params = ivt_load_default_run_params(cfg),
      best_row = NULL,
      source_tsv = NA_character_,
      load_error = conditionMessage(e)
    )
  }
)
run_params <- best_fit$run_params

lineage_run <- ivt_run_lineage(
  adapter = adapter,
  cfg = cfg,
  run_params = run_params
)

summary_df <- ivt_collect_lineage_summary(
  run = lineage_run,
  fit_data = ivt_objects$fit_data
)

dist_df <- ivt_collect_distribution_summary(lineage_run)
daily_count_df <- ivt_collect_daily_counts(lineage_run)
best_comp <- ivt_objective_components(
  run_params = run_params,
  fit_objects = ivt_objects,
  cfg = cfg,
  fallback_max_passage_days = fallback_max_passage_days,
  growth_weight = 1,
  ploidy_weight = 1
)

path_map <- bind_rows(
  ivt_terminal_path_map(ivt_objects$jobs_2N, cohort = "2N"),
  ivt_terminal_path_map(ivt_objects$jobs_4N, cohort = "4N")
)

objective_summary_df <- best_comp$summary |>
  left_join(path_map, by = c("cohort", "segment_id"), relationship = "many-to-many")

objective_dist_df <- bind_rows(
  ivt_collect_distribution_summary(best_comp$run_2N),
  ivt_collect_distribution_summary(best_comp$run_4N)
) |>
  left_join(path_map, by = c("cohort", "segment_id"), relationship = "many-to-many")
```

```{r best-fit-note, results='asis'}
if (!is.null(best_fit$best_row)) {
  cat(
    paste0(
      "Using best optimized parameter set from `", best_fit$source_tsv, "`  \n",
      "draw_id = `", best_fit$best_row$draw_id[[1]], "`  \n",
      "objective = `", signif(as.numeric(best_fit$best_row$objective[[1]]), 6), "`  \n",
      "terminal lineages in plotted objective = `", n_distinct(path_map$cohort, path_map$lineage_terminal_key), "`"
    )
  )
} else {
  load_error_msg <- if ("load_error" %in% names(best_fit) && nzchar(best_fit$load_error)) {
    best_fit$load_error
  } else {
    "unknown"
  }
  cat(
    paste0(
      "No sampled-objective TSV found. Using default parameter table values.  \n",
      "load error = `", load_error_msg, "`"
    )
  )
}
```

```{r daily-counts}
ivt_plot_daily_counts(daily_count_df)
```

```{r selected-counts}
ivt_plot_lineage_counts(summary_df)
```

```{r growth-rate}
ivt_plot_lineage_growth(objective_summary_df)
```

```{r ploidy-mean}
ivt_plot_lineage_ploidy(objective_summary_df)
```

```{r ploidy-distribution}
ivt_plot_distribution_heatmap(objective_dist_df, max_N = 110)
```
