---
title: "In Vitro Optimization Overview"
output:
  html_document:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 9.6,
  fig.height = 6,
  warning = FALSE,
  message = FALSE,
  echo=FALSE
)
```

```{r load}
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)

theme_invitro_report <- function() {
  ggplot2::theme_minimal(base_size = 8.5) +
    ggplot2::theme(
      plot.title = ggplot2::element_text(size = 7, face = "bold"),
      plot.subtitle = ggplot2::element_text(size = 7),
      axis.title = ggplot2::element_text(size = 7),
      axis.text = ggplot2::element_text(size = 7),
      legend.title = ggplot2::element_text(size = 7),
      legend.text = ggplot2::element_text(size = 7),
      strip.text = ggplot2::element_text(size = 7),
      plot.caption = ggplot2::element_text(size = 7)
    )
}

theme_set(theme_invitro_report())

repo_root <- local({
  candidates <- unique(normalizePath(
    c(getwd(), file.path(getwd(), ".."), file.path(getwd(), "..", "..")),
    mustWork = FALSE
  ))
  hit <- candidates[
    file.exists(file.path(candidates, "code", "in-vitro-utils", "io.R"))
  ][1]
  if (is.na(hit) || !nzchar(hit)) {
    stop("Could not locate repository root from working directory: ", getwd())
  }
  hit
})

source(file.path(repo_root, "code", "in-vitro-utils", "io.R"))
source(file.path(repo_root, "code", "in-vitro-utils", "lineage_adapter.R"))
source(file.path(repo_root, "code", "in-vitro-utils", "runner.R"))
source(file.path(repo_root, "code", "in-vitro-utils", "summaries.R"))
source(file.path(repo_root, "code", "in-vitro-utils", "plotting.R"))
source(file.path(repo_root, "code", "in-vitro-utils", "objective.R"))

ivt_source_map_model(repo_root)

inspection_top_n <- 100L
inspection_top_n <- max(1L, as.integer(inspection_top_n))
identifiability_step_frac <- 0.02
identifiability_eigen_count <- 5L
identifiability_loading_count <- 5L
```

This document reports the current fit of the oxygen/ploidy model to the in vitro lineage data.
The underlying in vitro data assembly and its derivation are documented in the `ploidyOxygen/` subfolder; this report focuses on the fitted parameter set, the optimization diagnostics, and the resulting model behaviour.

What it does:

- loads the precomputed in vitro fit objects (`fit_data.Rds`, `jobs_2N.Rds`, `jobs_4N.Rds`)
- loads the best optimized parameter set from `workflow/sampled_objective_draws`, falling back to the committed checkpoint file `data/O2_supply_demand/invitro_best_parameters.tsv` when sampled draws are absent
- evaluates the fitted objective across all in vitro 2N and 4N lineage segments
- summarizes optimizer behaviour across sampled-start local runs
- shows key model-internal rate functions at the optimized parameter values
- demonstrates how the existing simulation machinery is used to mimic the in vitro passaging experiments
- reports the resulting fit to growth and chromosome-count observables

```{r build-adapter}
ivt_objects <- ivt_load_fit_objects(repo_root)
fallback_max_passage_days <- 14

terminal_key <- ivt_choose_demo_terminal_key(
  jobs = ivt_objects$jobs_2N,
  fit_data = ivt_objects$fit_data,
  min_kary = 1L
)

adapter <- ivt_build_lineage_adapter(
  jobs = ivt_objects$jobs_2N,
  fit_data = ivt_objects$fit_data,
  terminal_sim_key = terminal_key,
  cohort = "2N",
  max_segment_days = fallback_max_passage_days
)
```

Chosen lineage: `r terminal_key`

Assumptions used here:

- one `jobs_2N` row is treated as one simulation segment
- if multiple `data_ids` attach to one row, `passage_duration` is aggregated by median
- current checkpoint simplification: segment `initial_cells` and `final_cells` are pooled to lineage/cohort-level medians rather than taken passage-by-passage
- fallback maximum passage length if duration is missing = `r fallback_max_passage_days` days

```{r run-lineage}
cfg <- ivt_build_default_cfg(
  repo_root = repo_root,
  dt = 0.1,
  init_total_size = 1e6,
  o2_upper_bound = 21,
  fixed_oxygen = TRUE
)

best_sample_path <- file.path(repo_root, "workflow", "sampled_objective_draws", "invitro_objective_array.tsv")
best_fit <- tryCatch(
  ivt_load_best_sampled_run_params(cfg = cfg, repo_root = repo_root, tsv_path = best_sample_path),
  error = function(e) {
    list(
      run_params = ivt_load_default_run_params(cfg),
      best_row = NULL,
      source_tsv = NA_character_,
      load_error = conditionMessage(e)
    )
  }
)
run_params <- best_fit$run_params

lineage_run <- ivt_run_lineage(
  adapter = adapter,
  cfg = cfg,
  run_params = run_params
)

summary_df <- ivt_collect_lineage_summary(
  run = lineage_run,
  fit_data = ivt_objects$fit_data
)

dist_df <- ivt_collect_distribution_summary(lineage_run)
daily_count_df <- ivt_collect_daily_counts(lineage_run)
best_comp <- ivt_objective_components(
  run_params = run_params,
  fit_objects = ivt_objects,
  cfg = cfg,
  fallback_max_passage_days = fallback_max_passage_days,
  growth_weight = 1,
  ploidy_weight = 1
)

path_map <- bind_rows(
  ivt_terminal_path_map(ivt_objects$jobs_2N, cohort = "2N"),
  ivt_terminal_path_map(ivt_objects$jobs_4N, cohort = "4N")
)

objective_summary_df <- best_comp$summary |>
  left_join(path_map, by = c("cohort", "segment_id"), relationship = "many-to-many")

objective_dist_df <- bind_rows(
  ivt_collect_distribution_summary(best_comp$run_2N),
  ivt_collect_distribution_summary(best_comp$run_4N)
) |>
  left_join(path_map, by = c("cohort", "segment_id"), relationship = "many-to-many")

lineage_alias <- if ("lineage_label" %in% names(summary_df)) {
  alias_vals <- unique(summary_df$lineage_label[is.finite(match(summary_df$lineage_label, summary_df$lineage_label))])
  alias_vals <- alias_vals[!is.na(alias_vals) & nzchar(alias_vals)]
  if (length(alias_vals) > 0L) alias_vals[[1]] else terminal_key
} else {
  terminal_key
}
```

```{r best-fit-note, results='asis'}
if (!is.null(best_fit$best_row)) {
  source_label <- if (identical(best_fit$source_type, "committed_best_parameters")) {
    "Using committed checkpoint parameter set from `"
  } else {
    "Using best optimized parameter set from `"
  }
  cat(
    paste0(
      source_label, best_fit$source_tsv, "`  \n",
      "optimization command = `sbatch workflow/submit_sample_invitro_objective_array.slurm`  \n",
      "draw_id = `", best_fit$best_row$draw_id[[1]], "`  \n",
      "objective = `", signif(as.numeric(best_fit$best_row$objective[[1]]), 6), "`  \n",
      "terminal lineages in plotted objective = `", n_distinct(path_map$cohort, path_map$lineage_terminal_key), "`"
    )
  )
} else {
  load_error_msg <- if ("load_error" %in% names(best_fit) && nzchar(best_fit$load_error)) {
    best_fit$load_error
  } else {
    "unknown"
  }
  cat(
    paste0(
      "Could not load sampled-objective draws or committed checkpoint parameters. Using default parameter table values.  \n",
      "load error = `", load_error_msg, "`"
    )
  )
}
```

```{r optimization-diagnostics-data}
sampled_diag <- tryCatch({
  sampled_tsv <- ivt_resolve_sampled_objective_tsv(
    repo_root = repo_root,
    tsv_path = best_sample_path
  )
  sampled_tab <- read.delim(sampled_tsv, stringsAsFactors = FALSE, check.names = FALSE)
  if (nrow(sampled_tab) == 0L || !"objective" %in% names(sampled_tab)) {
    stop("Sampled objective table is empty or missing objective column.")
  }

  sampled_tab$objective <- as.numeric(sampled_tab$objective)
  sampled_tab <- sampled_tab |>
    filter(is.finite(objective)) |>
    arrange(objective) |>
    mutate(
      run_rank = row_number(),
      objective_gap = objective - min(objective, na.rm = TRUE)
    )

  tol_grid <- c(1e-8, 1e-6, 1e-4, 1e-3)
  opt_counts <- vapply(
    tol_grid,
    function(tol) sum(sampled_tab$objective_gap <= tol, na.rm = TRUE),
    numeric(1)
  )
  optimum_note <- paste(
    paste0("within ", format(tol_grid, scientific = TRUE), ": n=", opt_counts),
    collapse = "; "
  )

  optim_spec <- ivt_optimizer_spec(cfg)
  transformed_cols <- intersect(optim_spec$param_name, names(sampled_tab))
  natural_spec <- read.csv(cfg$parameter_table, stringsAsFactors = FALSE) |>
    filter(tolower(trimws(as.character(use_invitro_fit))) %in% c("true", "t", "1", "yes", "y")) |>
    transmute(
      param_symbol = as.character(param_symbol),
      lower_bound = as.numeric(lower_bound),
      upper_bound = as.numeric(upper_bound)
    )
  param_name_map <- tibble(
    optim_param_symbol = c(
      "log10_lam_min", "log10_lam_max", "log10_k_o", "log10_p_misseg",
      "log10_k_o_mis", "log10_gamma_loss", "log10_p_wgd",
      "log10_alpha_o2", "gamma_growth", "log10_mu_hp",
      "gamma_mu", "log10_O2_crit", "n_O", "log10_p_mis_base"
    ),
    param_symbol = c(
      "lam_min", "lam_max", "k_o", "p_misseg",
      "k_o_mis", "gamma_loss", "p_wgd",
      "alpha_o2", "gamma_growth", "mu_hp",
      "gamma_mu", "O2_crit", "n_O", "p_mis_base"
    )
  )
  top_n <- min(inspection_top_n, nrow(sampled_tab))
  top_runs <- sampled_tab |>
    slice_head(n = top_n)

  compare_rank <- top_n
  compare_row <- top_runs |>
    slice(compare_rank)

  parameter_bounds_long <- if (length(transformed_cols) > 0L) {
    top_runs |>
      select(run_rank, draw_id, objective, all_of(transformed_cols)) |>
      pivot_longer(
        cols = all_of(transformed_cols),
        names_to = "column_name",
        values_to = "value"
      ) |>
      mutate(
        param_symbol = column_name,
        value = as.numeric(value)
      ) |>
      left_join(
        optim_spec |>
          transmute(
            param_symbol = param_name,
            lower_bound = as.numeric(lower),
            upper_bound = as.numeric(upper)
          ),
        by = "param_symbol"
      ) |>
      filter(is.finite(value), is.finite(lower_bound), is.finite(upper_bound), upper_bound > lower_bound) |>
      mutate(
        bound_fraction = (value - lower_bound) / (upper_bound - lower_bound),
        bound_status = case_when(
          bound_fraction <= 0.02 ~ "Near lower bound",
          bound_fraction >= 0.98 ~ "Near upper bound",
          TRUE ~ "Interior"
        ),
        param_symbol = factor(param_symbol, levels = unique(param_symbol))
      )
  } else {
    NULL
  }

  parameter_values_long <- if (length(transformed_cols) > 0L && nrow(natural_spec) > 0L) {
    bound_status_lookup <- parameter_bounds_long |>
      transmute(
        run_rank,
        draw_id,
        objective,
        optim_param_symbol = as.character(param_symbol),
        bound_status
      ) |>
      left_join(param_name_map, by = "optim_param_symbol") |>
      filter(!is.na(param_symbol)) |>
      select(run_rank, draw_id, objective, param_symbol, bound_status)

    bind_rows(lapply(seq_len(nrow(top_runs)), function(i) {
      run_i <- top_runs[i, , drop = FALSE]
      par_t_i <- suppressWarnings(as.numeric(run_i[1, transformed_cols, drop = TRUE]))
      names(par_t_i) <- transformed_cols
      run_params_i <- ivt_optim_par_to_run_params(par_t = par_t_i, cfg = cfg)

      tibble(
        run_rank = run_i$run_rank[[1]],
        draw_id = run_i$draw_id[[1]],
        objective = run_i$objective[[1]],
        param_symbol = natural_spec$param_symbol,
        value = vapply(natural_spec$param_symbol, function(nm) as.numeric(run_params_i[[nm]]), numeric(1))
      )
    })) |>
      left_join(
        bound_status_lookup,
        by = c("run_rank", "draw_id", "objective", "param_symbol")
      ) |>
      left_join(natural_spec, by = "param_symbol") |>
      filter(
        is.finite(value),
        is.finite(lower_bound),
        is.finite(upper_bound),
        lower_bound > 0,
        upper_bound > lower_bound
      ) |>
      mutate(
        param_symbol = factor(param_symbol, levels = natural_spec$param_symbol)
      )
  } else {
    NULL
  }

  convergence_tab <- if ("convergence" %in% names(sampled_tab)) {
    sampled_tab |>
      mutate(
        convergence = as.integer(convergence),
        convergence_label = ifelse(
          is.na(convergence),
          "NA",
          ifelse(convergence == 0L, "0 (converged)", paste0(convergence))
        ),
        message_label = if ("message" %in% names(sampled_tab)) {
          ifelse(is.na(message) | !nzchar(trimws(message)), "No message", trimws(message))
        } else {
          "No message column"
        }
      )
  } else {
    NULL
  }

  list(
    source_tsv = sampled_tsv,
    sampled_tab = sampled_tab,
    optimum_note = optimum_note,
    top_n = top_n,
    top_runs = top_runs,
    compare_rank = compare_rank,
    compare_row = compare_row,
    parameter_bounds_long = parameter_bounds_long,
    parameter_values_long = parameter_values_long,
    convergence_tab = convergence_tab
  )
}, error = function(e) {
  list(
    error = conditionMessage(e),
    sampled_tab = NULL,
    parameter_bounds_long = NULL,
    parameter_values_long = NULL,
    convergence_tab = NULL
  )
})
```

## Optimization diagnostics

These diagnostics summarize the sampled-start local optimization runs. For the objective-value panel, the main stability signal is many runs converging to similar objective values. The box-bounds and natural-scale parameter panels show whether the best runs are consistently pushing against fitted parameter bounds, while the convergence-code summary shows whether the optimizer terminated cleanly.

```{r optimization-diagnostics-note, results='asis'}
if (!is.null(sampled_diag$sampled_tab)) {
  cat(
    paste0(
      "Diagnostics source TSV: `", sampled_diag$source_tsv, "`  \n",
      "runs loaded = `", nrow(sampled_diag$sampled_tab), "`  \n",
      sampled_diag$optimum_note
    )
  )
} else {
  if (identical(best_fit$source_type, "committed_best_parameters")) {
    cat(
      paste0(
        "Optimizer diagnostics unavailable.  \n",
        "No sampled-objective draw tables were found, so this report is using the committed checkpoint parameter file instead."
      )
    )
  } else {
    cat(
      paste0(
        "Optimizer diagnostics unavailable.  \n",
        "load error = `", sampled_diag$error, "`"
      )
    )
  }
}
```

```{r optimization-diagnostic-panels, fig.height=10.5}
if (!is.null(sampled_diag$sampled_tab)) {
  gap_eps <- max(min(sampled_diag$sampled_tab$objective_gap[sampled_diag$sampled_tab$objective_gap > 0], na.rm = TRUE), 1e-12)
  if (!is.finite(gap_eps)) gap_eps <- 1e-12

  p_optimum <- sampled_diag$sampled_tab |>
    mutate(top_group = ifelse(run_rank <= sampled_diag$top_n, paste0("Top ", sampled_diag$top_n), "Other runs")) |>
    ggplot(aes(run_rank, objective_gap + gap_eps, color = top_group)) +
    geom_point(alpha = 0.8, size = 2) +
    scale_color_manual(values = stats::setNames(c("#e15759", "#4c78a8"), c(paste0("Top ", sampled_diag$top_n), "Other runs"))) +
    scale_y_log10() +
    labs(
      title = "Optimizer objective values across runs",
      subtitle = paste0("Highlighted points mark the top ", sampled_diag$top_n, " runs carried forward for inspection"),
      x = "Optimization run rank",
      y = "Objective gap from best run (log10 scale)",
      color = NULL
    )

  p_bounds <- if (!is.null(sampled_diag$parameter_bounds_long) && nrow(sampled_diag$parameter_bounds_long) > 0L) {
    sampled_diag$parameter_bounds_long |>
    ggplot(aes(bound_fraction, param_symbol, color = bound_status)) +
    geom_vline(xintercept = c(0, 1), linetype = "dashed", color = "grey70") +
    geom_point(aes(size = -log10(pmax(objective, 1e-12))), alpha = 0.8, position = position_jitter(height = 0.15, width = 0)) +
    scale_color_manual(
      values = c(
        "Near lower bound" = "#c44e52",
        "Interior" = "#4c78a8",
        "Near upper bound" = "#f58518"
      )
    ) +
    labs(
      title = paste0("Top ", sampled_diag$top_n, " runs: optimizer parameters relative to box bounds"),
      subtitle = NULL,
      x = "Position within optimizer box bounds",
      y = NULL,
      color = "Bound status",
      size = "-log10(objective)"
    )
  } else {
    ggplot() + theme_void() + labs(title = "No bound diagnostics available")
  }

  p_values <- if (!is.null(sampled_diag$parameter_values_long) && nrow(sampled_diag$parameter_values_long) > 0L) {
    sampled_diag$parameter_values_long |>
      ggplot(aes(value, param_symbol, color = bound_status)) +
      geom_segment(
        data = distinct(sampled_diag$parameter_values_long, param_symbol, lower_bound, upper_bound),
        aes(x = lower_bound, xend = upper_bound, y = param_symbol, yend = param_symbol),
        inherit.aes = FALSE,
        linewidth = 0.5,
        color = "grey70"
      ) +
      geom_point(aes(size = -log10(pmax(objective, 1e-12))), alpha = 0.8, position = position_jitter(height = 0.15, width = 0)) +
      scale_x_log10() +
      scale_color_manual(
        values = c(
          "Near lower bound" = "#c44e52",
          "Interior" = "#4c78a8",
          "Near upper bound" = "#f58518"
        )
      ) +
      labs(
        title = paste0("Top ", sampled_diag$top_n, " runs: fitted parameter values"),
        subtitle = NULL,
        x = "Parameter value (log10 scale)",
        y = NULL,
        color = "Bound status",
        size = "-log10(objective)"
      )
  } else {
    ggplot() + theme_void() + labs(title = "No fitted-parameter diagnostics available")
  }

  p_conv <- if (!is.null(sampled_diag$convergence_tab)) {
    sampled_diag$convergence_tab |>
    count(convergence_label, sort = TRUE) |>
    ggplot(aes(reorder(convergence_label, n), n, fill = convergence_label == "0 (converged)")) +
    geom_col(show.legend = FALSE) +
    coord_flip() +
    scale_fill_manual(values = c("TRUE" = "#59a14f", "FALSE" = "#e15759")) +
    labs(
      title = "Recorded optimizer convergence codes",
      subtitle = NULL,
      x = "Convergence code",
      y = "Number of runs"
    )
  } else {
    ggplot() + theme_void() + labs(title = "No convergence codes recorded")
  }

  top_row <- gridExtra::arrangeGrob(p_optimum, p_conv, nrow = 1)
  gridExtra::grid.arrange(top_row, p_bounds, p_values, nrow = 3, heights = c(1, 1.2, 1.2))
}
```

```{r top-n-comparison-data}
comparison_fit <- NULL
comparison_summary_df <- NULL
comparison_dist_df <- NULL
comparison_label <- NULL

if (!is.null(sampled_diag$compare_row) && nrow(sampled_diag$compare_row) == 1L) {
  compare_row <- sampled_diag$compare_row
  compare_transformed <- intersect(ivt_optimizer_spec(cfg)$param_name, names(compare_row))
  compare_par_t <- suppressWarnings(as.numeric(compare_row[1, compare_transformed, drop = TRUE]))
  names(compare_par_t) <- compare_transformed
  compare_run_params <- ivt_optim_par_to_run_params(par_t = compare_par_t, cfg = cfg)
  comparison_fit <- ivt_objective_components(
    run_params = compare_run_params,
    fit_objects = ivt_objects,
    cfg = cfg,
    fallback_max_passage_days = fallback_max_passage_days,
    growth_weight = 1,
    ploidy_weight = 1
  )
  comparison_summary_df <- comparison_fit$summary |>
    left_join(path_map, by = c("cohort", "segment_id"), relationship = "many-to-many")
  comparison_dist_df <- bind_rows(
    ivt_collect_distribution_summary(comparison_fit$run_2N),
    ivt_collect_distribution_summary(comparison_fit$run_4N)
  ) |>
    left_join(path_map, by = c("cohort", "segment_id"), relationship = "many-to-many")
  comparison_label <- paste0("Rank ", sampled_diag$compare_rank)
}
```

## Identifiability diagnostics

This section adds a cheap practical identifiability check around the best recovered parameter set. It uses a finite-difference Jacobian of the weighted residual vector in optimizer-transformed parameter space, forms `J^T J`, and reports local conditioning diagnostics. In this transformed space, most fitted parameters are on a `log10` scale; `gamma_growth`, `gamma_mu`, and `n_O` remain untransformed.

```{r identifiability-data}
identifiability_diag <- tryCatch({
  optim_spec <- ivt_optimizer_spec(cfg)
  best_par_t <- ivt_run_params_to_optim_par(run_params, cfg = cfg)
  best_par_t <- as.numeric(best_par_t)
  names(best_par_t) <- optim_spec$param_name

  ivt_weighted_residual_vector <- function(par_t) {
    run_params_i <- ivt_optim_par_to_run_params(par_t = par_t, cfg = cfg)
    comp_i <- ivt_objective_components(
      run_params = run_params_i,
      fit_objects = ivt_objects,
      cfg = cfg,
      fallback_max_passage_days = fallback_max_passage_days,
      growth_weight = 1,
      ploidy_weight = 1
    )
    summary_i <- comp_i$summary
    growth_df_i <- summary_i |>
      filter(is.finite(observed_growth), is.finite(predicted_growth_rate))
    kary_df_i <- summary_i |>
      filter(is.finite(observed_mean_kary_N), is.finite(predicted_mean_kary_N))

    growth_scale <- if (nrow(growth_df_i) > 0L) sqrt(1 / nrow(growth_df_i)) else NULL
    kary_scale <- if (nrow(kary_df_i) > 0L) sqrt(1 / nrow(kary_df_i)) else NULL

    c(
      if (!is.null(growth_scale)) growth_scale * (growth_df_i$predicted_growth_rate - growth_df_i$observed_growth),
      if (!is.null(kary_scale)) kary_scale * (kary_df_i$predicted_mean_kary_N - kary_df_i$observed_mean_kary_N)
    )
  }

  residual_best <- ivt_weighted_residual_vector(best_par_t)
  if (!length(residual_best)) stop("Residual vector is empty at the best fit.")

  step_sizes <- pmax(
    identifiability_step_frac * (optim_spec$upper - optim_spec$lower),
    1e-4
  )
  J <- matrix(NA_real_, nrow = length(residual_best), ncol = length(best_par_t))
  colnames(J) <- names(best_par_t)

  for (j in seq_along(best_par_t)) {
    step_j <- step_sizes[[j]]
    par_minus <- best_par_t
    par_plus <- best_par_t
    par_minus[[j]] <- max(optim_spec$lower[[j]], best_par_t[[j]] - step_j)
    par_plus[[j]] <- min(optim_spec$upper[[j]], best_par_t[[j]] + step_j)

    actual_step <- par_plus[[j]] - par_minus[[j]]
    if (!is.finite(actual_step) || actual_step <= 0) {
      next
    }

    resid_minus <- ivt_weighted_residual_vector(par_minus)
    resid_plus <- ivt_weighted_residual_vector(par_plus)
    if (length(resid_minus) != length(residual_best) || length(resid_plus) != length(residual_best)) {
      stop("Residual vector length changed during finite-difference evaluation.")
    }
    J[, j] <- (resid_plus - resid_minus) / actual_step
  }

  valid_cols <- which(colSums(is.finite(J)) == nrow(J))
  if (!length(valid_cols)) stop("No finite Jacobian columns were available.")
  J <- J[, valid_cols, drop = FALSE]

  jt_j <- crossprod(J)
  eig <- eigen(jt_j, symmetric = TRUE)
  eig_values <- as.numeric(eig$values)
  eig_vectors <- eig$vectors
  rownames(eig_vectors) <- colnames(J)
  positive_eigs <- eig_values[eig_values > .Machine$double.eps]
  condition_number <- if (length(positive_eigs) >= 2L) max(positive_eigs) / min(positive_eigs) else Inf

  eigen_count <- min(identifiability_eigen_count, length(eig_values))
  weakest_idx <- seq.int(
    from = max(1L, length(eig_values) - eigen_count + 1L),
    to = length(eig_values)
  )
  weakest_eigs <- tibble(
    eigen_rank = seq_len(eigen_count),
    eigenvalue = eig_values[weakest_idx],
    reciprocal_sqrt = ifelse(eig_values[weakest_idx] > 0, 1 / sqrt(eig_values[weakest_idx]), Inf)
  )

  weakest_loadings <- bind_rows(lapply(seq_along(weakest_idx), function(i) {
    idx <- weakest_idx[[i]]
    tibble(
      eigen_rank = i,
      param_symbol = rownames(eig_vectors),
      loading = eig_vectors[, idx]
    ) |>
      mutate(abs_loading = abs(loading)) |>
      arrange(desc(abs_loading)) |>
      slice_head(n = identifiability_loading_count)
  }))

  cor_try <- tryCatch({
    cov_mat <- solve(jt_j)
    sd_vec <- sqrt(diag(cov_mat))
    cor_mat <- cov_mat / outer(sd_vec, sd_vec)
    cor_mat[!is.finite(cor_mat)] <- NA_real_
    cor_mat
  }, error = function(e) NULL)

  cor_long <- if (!is.null(cor_try)) {
    cor_df <- as.data.frame(as.table(cor_try), stringsAsFactors = FALSE)
    names(cor_df) <- c("param_1", "param_2", "correlation")
    cor_df |>
      filter(param_1 != param_2, is.finite(correlation))
  } else {
    NULL
  }

  list(
    residual_length = length(residual_best),
    jacobian = J,
    jt_j = jt_j,
    condition_number = condition_number,
    weakest_eigs = weakest_eigs,
    weakest_loadings = weakest_loadings,
    correlation_matrix = cor_try,
    correlation_long = cor_long
  )
}, error = function(e) {
  list(error = conditionMessage(e))
})
```

```{r identifiability-note, results='asis'}
if (is.null(identifiability_diag$error)) {
  cat(
    paste0(
      "Residual length = `", identifiability_diag$residual_length, "`  \n",
      "Jacobian columns = `", ncol(identifiability_diag$jacobian), "`  \n",
      "Finite-difference step fraction of parameter-box width = `", signif(identifiability_step_frac, 3), "`  \n",
      "Condition number of `J^T J` = `", signif(identifiability_diag$condition_number, 4), "`"
    )
  )
} else {
  cat(
    paste0(
      "Identifiability diagnostics unavailable.  \n",
      "load error = `", identifiability_diag$error, "`"
    )
  )
}
```

```{r identifiability-panels, fig.height=8.5}
if (is.null(identifiability_diag$error)) {
  p_eigs <- identifiability_diag$weakest_eigs |>
    ggplot(aes(reorder(paste0("Weakest ", eigen_rank), eigenvalue), eigenvalue)) +
    geom_col(fill = "#4e79a7") +
    scale_y_log10() +
    coord_flip() +
    labs(
      title = "Smallest eigenvalues of J^T J",
      x = NULL,
      y = "Eigenvalue (log10 scale)"
    )

  p_loadings <- identifiability_diag$weakest_loadings |>
    mutate(
      eigen_label = paste0("Weakest ", eigen_rank),
      param_symbol = reorder(paste0(param_symbol), abs_loading)
    ) |>
    ggplot(aes(abs_loading, param_symbol, fill = loading > 0)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~ eigen_label, scales = "free_y") +
    scale_fill_manual(values = c("TRUE" = "#59a14f", "FALSE" = "#e15759")) +
    labs(
      title = "Dominant parameter loadings in weakest eigenvectors",
      x = "Absolute loading",
      y = NULL
    )

  p_corr <- if (!is.null(identifiability_diag$correlation_long) && nrow(identifiability_diag$correlation_long) > 0L) {
    identifiability_diag$correlation_long |>
      mutate(
        param_1 = factor(param_1, levels = colnames(identifiability_diag$jacobian)),
        param_2 = factor(param_2, levels = colnames(identifiability_diag$jacobian))
      ) |>
      ggplot(aes(param_1, param_2, fill = correlation)) +
      geom_tile() +
      scale_fill_gradient2(low = "#4e79a7", mid = "white", high = "#e15759", limits = c(-1, 1)) +
      labs(
        title = "Parameter correlation matrix from (J^T J)^-1",
        x = NULL,
        y = NULL,
        fill = "Correlation"
      ) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  } else {
    ggplot() + theme_void() + labs(title = "Correlation matrix unavailable")
  }

  gridExtra::grid.arrange(p_eigs, p_loadings, p_corr, nrow = 3, heights = c(0.8, 1.3, 1.6))
}
```

```{r identifiability-tables}
if (is.null(identifiability_diag$error)) {
  list(
    weakest_eigenvalues = identifiability_diag$weakest_eigs,
    weakest_eigenvector_loadings = identifiability_diag$weakest_loadings,
    strongest_parameter_correlations = if (!is.null(identifiability_diag$correlation_long)) {
      identifiability_diag$correlation_long |>
        mutate(abs_correlation = abs(correlation)) |>
        filter(as.character(param_1) < as.character(param_2)) |>
        arrange(desc(abs_correlation)) |>
        slice_head(n = 15) |>
        select(param_1, param_2, correlation)
    } else {
      NULL
    }
  )
}
```

```{r convergence-message-table}
if (!is.null(sampled_diag$convergence_tab) && "message_label" %in% names(sampled_diag$convergence_tab)) {
  sampled_diag$convergence_tab |>
    count(convergence_label, message_label, sort = TRUE) |>
    arrange(match(convergence_label, unique(sampled_diag$convergence_tab$convergence_label)), desc(n)) |>
    mutate(message_label = ifelse(nchar(message_label) > 90, paste0(substr(message_label, 1, 87), "..."), message_label))
}
```

## Rate-function diagnostics

The in vivo model uses cell-count dependent oxygen dynamics. In the in vitro workflow we use the existing oxygen-handling functions in an externally driven way, fixing the passage oxygen condition and using this plot as a diagnostic for how that reused machinery behaves in the current fit.

```{r rate-function-data}
o2_grid <- seq(0, 21, length.out = 300)
burden_grid <- seq(0, 5 * cfg$o2_Nref, length.out = 300)
ploidy_levels <- c(44, 66, 88, 110)
loss_sizes <- c(1, 2, 4)

hypoxia_weight_df <- tibble(
  O2 = o2_grid,
  hypoxia_weight = (pmax(run_params$O2_crit, 1e-12)^run_params$n_O) /
    ((pmax(run_params$O2_crit, 1e-12)^run_params$n_O) + (pmax(o2_grid, 0)^run_params$n_O))
)

oxygen_levels_used <- sort(unique(objective_summary_df$oxygen_pct[is.finite(objective_summary_df$oxygen_pct)]))
if (length(oxygen_levels_used) == 0L) {
  oxygen_levels_used <- sort(unique(c(adapter$segments[[1]]$oxygen_pct)))
}
oxygen_supply_df <- bind_rows(lapply(
  oxygen_levels_used,
  function(target_o2) {
    ivt_build_segment_o2_profile(
      target_o2_pct = target_o2,
      cfg = cfg,
      run_params = run_params,
      burden_grid = burden_grid
    )
  }
))

growth_base_df <- tibble(
  O2 = o2_grid,
  lambda_base = growth_lambda(
    O2 = o2_grid,
    N = 44,
    lam_min = run_params$lam_min,
    lam_max = run_params$lam_max,
    k_o = run_params$k_o
  )
)

growth_eff_df <- tidyr::expand_grid(
  O2 = o2_grid,
  N = ploidy_levels
) |>
  mutate(
    lambda_eff = vapply(
      seq_len(n()),
      function(i) {
        N_i <- N[[i]]
        O2_i <- O2[[i]]
        lam_base_i <- growth_lambda(
          O2 = O2_i,
          N = N_i,
          lam_min = run_params$lam_min,
          lam_max = run_params$lam_max,
          k_o = run_params$k_o
        )[[1]]
        h_o2_i <- hypoxia_weight_df$hypoxia_weight[which.min(abs(hypoxia_weight_df$O2 - O2_i))]
        lam_base_i / (1 + run_params$alpha_o2 * h_o2_i * ((N_i / 44)^run_params$gamma_growth))
      },
      numeric(1)
    ),
    N_label = paste0("N=", N)
  )

mu_eff_df <- tidyr::expand_grid(
  O2 = o2_grid,
  N = ploidy_levels
) |>
  mutate(
    mu_eff = .mu_eff_of_O2(O2 = O2, run_params = run_params, N = N),
    N_label = paste0("N=", N)
  )

pmisseg_df <- tidyr::expand_grid(
  O2 = o2_grid,
  N = ploidy_levels
) |>
  mutate(
    p_misseg_eff = .pmisseg_of_O2(O2 = O2, run_params = run_params, N = N),
    N_label = paste0("N=", N)
  )

loss_survival_df <- tidyr::expand_grid(
  N = seq(44, min(cfg$N_MAX, 154), by = 2),
  m_loss = loss_sizes
) |>
  mutate(
    survival = mapply(
      function(N_i, m_i) {
        .loss_survival_nullisomy(
          N = N_i,
          m_loss = m_i,
          gamma_loss = run_params$gamma_loss,
          N_unit = cfg$N_UNIT
        )[[1]]
      },
      N,
      m_loss
    ),
    loss_label = paste0("m_loss=", m_loss)
  )

mu_title <- {
  death_mode <- as.character(run_params$ploidy_O2_death)
  if (identical(death_mode, "uniform")) {
    "mu_eff(O2) = mu_hp * h(O2)"
  } else if (identical(death_mode, "ploidy_related")) {
    "mu_eff(O2,N) = mu_hp * h(O2) * (N/44)^gamma_mu"
  } else {
    "mu_eff(O2,N) = mu_hp * h(O2) * (1 + max(N/44 - 1, 0)^gamma_mu)"
  }
}
```

```{r oxygen-supply-function, fig.width=6, fig.height=4}
if (isTRUE(cfg$o2_burden_feedback)) {
  ggplot(oxygen_supply_df, aes(burden, oxygen_target)) +
    geom_line(linewidth = 1, color = "#1f77b4") +
    labs(
      title = "Burden-coupled oxygen target under the supply-demand feedback law",
      subtitle = "O2(Ntot) = max(o2_min, o2_S0 - kappa_O * log(1 + Ntot / o2_Nref))",
      x = "Viable burden proxy Ntot",
      y = "Target oxygen (%)"
    )
} else {
  ggplot(oxygen_supply_df, aes(burden, oxygen_target, color = oxygen_mode)) +
    geom_line(linewidth = 1) +
    labs(
      title = "Constant external oxygen used by the in vitro runner for each passage condition",
      subtitle = "Fixed-O2 in vitro runner: ivt_set_segment_o2(target_O2) => o2_S0 = o2_min = target_O2",
      x = "Viable burden proxy Ntot",
      y = "Target oxygen (%)",
      color = "Passage oxygen"
    )
}
```

The six panels below show key model-internal functions evaluated at the optimized parameter values. Together they summarize how oxygen, ploidy, death, missegregation, and loss survival interact in the fitted system.

```{r rate-function-panels, fig.height=10}
p_hypoxia <- ggplot(hypoxia_weight_df, aes(O2, hypoxia_weight)) +
  geom_line(linewidth = 1, color = "#d62728") +
  labs(
    title = "Extent of hypoxia on a 0-1 scale used by growth and death modules",
    subtitle = "h(O2) = O2_crit^n_O / (O2_crit^n_O + O2^n_O)",
    x = "Oxygen (%)",
    y = "Hypoxia weight"
  )

p_growth_base <- ggplot(growth_base_df, aes(O2, lambda_base)) +
  geom_line(linewidth = 1, color = "#2ca02c") +
  labs(
    title = "Baseline oxygen-limited proliferation rate before ploidy penalty",
    subtitle = "lambda_base(O2) = lam_min + (lam_max - lam_min) * O2 / (O2 + k_o)",
    x = "Oxygen (%)",
    y = "Baseline growth rate"
  )

p_growth_eff <- ggplot(growth_eff_df, aes(O2, lambda_eff, color = N_label)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Oxygen & ploidy dependent proliferation rate",
    subtitle = "lambda_eff(O2,N) = lambda_base(O2) / (1 + alpha_o2 * h(O2) * (N/44)^gamma_growth)",
    x = "Oxygen (%)",
    y = "Effective growth rate",
    color = "Ploidy state"
  )

p_death <- ggplot(mu_eff_df, aes(O2, mu_eff, color = N_label)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Hypoxia-linked death hazard",
    subtitle = mu_title,
    x = "Oxygen (%)",
    y = "Hypoxia-linked death rate",
    color = "Ploidy state"
  )

p_misseg <- ggplot(pmisseg_df, aes(O2, p_misseg_eff, color = N_label)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Per-chromosome missegregation probability",
    subtitle = "p_mis(O2,N) = p_mis_base + p_misseg * mu_eff(O2,N) / (mu_eff(O2,N) + k_o_mis)",
    x = "Oxygen (%)",
    y = "Per-chromosome missegregation probability",
    color = "Ploidy state"
  )

p_loss <- ggplot(loss_survival_df, aes(N, survival, color = loss_label)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Survival modifier applied to loss-branch missegregation outcomes",
    subtitle = "s_loss(N,m) = (1 - R_null(N,m))^gamma_loss",
    x = "Mother ploidy state N",
    y = "Loss-branch survival modifier",
    color = "Loss size"
  )

gridExtra::grid.arrange(
  p_hypoxia,
  p_growth_base,
  p_growth_eff,
  p_death,
  p_misseg,
  p_loss,
  nrow = 3
)
```

The next two plots illustrate how we mimic the in vitro passaging experiment using the existing simulation code. For each passage we have an observed initial cell count and a target final cell count derived from the data. We simulate for the recorded passage duration, then pick the simulated time point whose live-cell count is closest to the target count; the chromosome-count distribution at that point is propagated forward to seed the next passage.

```{r daily-counts}
ivt_plot_daily_counts(daily_count_df) +
  labs(subtitle = paste0("Lineage shown: ", lineage_alias))
```

```{r selected-counts}
ivt_plot_lineage_counts(summary_df) +
  labs(subtitle = paste0("Lineage shown: ", lineage_alias))
```

The final three plots summarize the objective-function fit across all lineage segments. The objective combines a growth component and a mean-chromosome-count component, each evaluated as a mean squared error between the observed in vitro measurements and the corresponding model predictions.

```{r fit-comparison-note, results='asis'}
if (!is.null(comparison_summary_df) && !is.null(sampled_diag$compare_rank)) {
  cat(
    paste0(
      "For the growth-rate and mean-chromosome-count panels, the best recovered fit is overlaid with the rank-`",
      sampled_diag$compare_rank,
      "` fit among the top `",
      sampled_diag$top_n,
      "` sampled runs."
    )
  )
}
```

```{r growth-rate}
ivt_plot_lineage_growth(
  objective_summary_df,
  comparison_summary_df = comparison_summary_df,
  primary_label = "Best fit",
  comparison_label = comparison_label
)
```

```{r ploidy-mean}
ivt_plot_lineage_ploidy(
  objective_summary_df,
  comparison_summary_df = comparison_summary_df,
  primary_label = "Best fit",
  comparison_label = comparison_label
)
```

```{r ploidy-distribution}
ivt_plot_distribution_heatmap(objective_dist_df, max_N = 110)
```
